PyPy: raising the level for Dynamic Language implementations

### 2007-05-27 talk/dyla2007/dyla.tex
Swap the two paras of the abstract.  Experimental.

Dynamic languages are traditionally implemented by writing a virtual machine for
them, centered around an interpreter and/or a built-in compiler and providing
the object model and memory management. When a language becomes more successful,
the limitations of such an implementation lead to the emergence of alternative
implementations that try to solve some of the problems. Another reason for new
implementations is the desire to have the language integrate well with existing,
well-tuned object-oriented virtual machine like Java's. In this paper, we
describe the mechanisms that lead to an abundance of implementations, and
explore some of the limitations of standard VMs.  We propose a different
complementary alternative to writing VMs by hand and dealing with low-level
details: flexibly generating virtual machines from a single abstract language
``specification''.

This approach has been validated by the PyPy project that tries to explore
meta-programming facilities with which we can automatically insert many features
and low-level details into an automatically generated virtual machines --
including good just-in-time compilers tuned to the dynamic language at hand.
We thus argue in this paper that one should not write interpreters manually but
rather use meta-programming techniques and raise the overall level at which
dynamic languages are implemented. We believe this to be ultimately a better
investment of efforts than the development of more advanced general-purpose OO
VMs. \footnote{This research is partially supported by the EU funded
 project: IST 004779 PyPy (PyPy: Implementing Python in Python).} \\

>>>>

We argue in this paper that one should not write interpreters manually but
rather use meta-programming techniques and raise the overall level at which
dynamic languages are implemented. We believe this to be ultimately a better
investment of efforts than the development of more advanced general-purpose OO
VMs.

Dynamic languages are traditionally implemented by writing a virtual machine for
them, centered around an interpreter and/or a built-in compiler and providing
the object model and memory management. When a language becomes more successful,
the limitations of such an implementation lead to the emergence of alternative
implementations that try to solve some of the problems. Another reason for new
implementations is the desire to have the language integrate well with existing,
well-tuned object-oriented virtual machine like Java's. In this paper, we
describe the mechanisms that lead to an abundance of implementations, and
explore some of the limitations of standard VMs.  We propose a different
complementary alternative to writing VMs by hand and dealing with low-level
details, validated by the PyPy project: flexibly generating virtual machines
from a single abstract language ``specification'', inserting features and
low-level details automatically -- including good just-in-time compilers tuned
to the dynamic language at hand.
\footnote{This research is partially supported by the EU funded
 project: IST 004779 PyPy (PyPy: Implementing Python in Python).} \\
